---
title: "Wi-Fi Locationing"
author: "Murilo Miranda"
date: "3/30/2020"
output:
  html_document:
    toc: true # table of content true
    toc_float: true
    code_folding: hide
    toc_depth: 2  # upto three depths of headings (specified by #, ## and ###)
    theme: united
    highlight: breezedark  # specifies the syntax highlighting style
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE, message=FALSE}
library(here)
library(tictoc)
library(knitr)
library(scatterplot3d)
library(tidyverse)
library(caret)
library(e1071)
```

## Introduction

Our client is developing a system to be deployed on large industrial campuses, in shopping malls, et cetera to help people to navigate a complex, unfamiliar interior space without getting lost. While GPS works fairly reliably outdoors, it generally doesn't work indoors, so a different technology is necessary. Our client would like us to investigate the feasibility of using "wifi fingerprinting" to determine a person's location in indoor spaces. Wifi fingerprinting uses the signals from multiple wifi hotspots within the building to determine location, analogously to how GPS uses satellite signals. We have been provided with a large database of wifi fingerprints for a multi-building industrial campus with a location (building, floor, and location ID) associated with each fingerprint. 

The goal is to evaluate multiple machine learning models to see which produces the best result, enabling us to make a recommendation to the client.

## Load datasets
```{r}
trainingData <- read.csv(here("/data/trainingData.csv"))
validationData <- read.csv(here("/data/validationData.csv"))

str(trainingData[, c(1:3, 519:529)])
```

## Preprocessing

### Convert data types
```{r}
# convert features to numeric
trainingData <- sapply(trainingData, as.numeric)
  
trainingData <- as_tibble(trainingData)
```

### Impute NA's
```{r}
# convert NA's to -110
trainingData[trainingData == 100] <- -110
```

### Remove features and Rows with all NA's
```{r}
# remove features with all data is NA's
trainingData <- trainingData[, colSums(trainingData == -110) < nrow(trainingData)]

# remove rows with all data is NA's
trainingData <- trainingData[rowSums(trainingData == -110) != ncol(trainingData), ]
```

### Create unique identifier
```{r}
# consolidate position identifiers to create location ID feature
trainingData$ID <- trainingData %>%
  group_indices(BUILDINGID, FLOOR)
```

```{r}
# convert features to categorical
trainingData <- trainingData %>% 
  mutate(BUILDINGID = as.factor(BUILDINGID),
         SPACEID = as.factor(SPACEID),
         RELATIVEPOSITION = as.factor(RELATIVEPOSITION),
         FLOOR = as.factor(FLOOR),
         ID = as.factor(ID))
```

```{r}
# 3D image of reference point locations in data set
scatterplot3d(trainingData$LONGITUDE, trainingData$LATITUDE, trainingData$FLOOR,
              type = 'p',
              highlight.3d = FALSE,
              color = 'red',
              angle = 200,
              scale.y = 0.9,
              pch = 20,
              box = FALSE,
              xlab = 'Longitude', ylab='Latitude',zlab = 'Building Floor') 
```

## Training with different models
### Set some parameters up

```{r}
# define parameters in trainControl
fitControl <- trainControl(method = "cv", number = 10)
```

### Models
#### Classification
```{r}
# remove unneeded features prior to model fitting
partTrainingData <- trainingData %>% 
  select(starts_with("WAP"), ID)
```

```{r}
# Randomly select - sample fixed fraction per group
set.seed(1605)
partTrainingData <- partTrainingData %>% sample_frac(size = 0.1, weight = ID)
```

```{r}
#
set.seed(1605)
in_training <- createDataPartition(partTrainingData$ID, p = 0.7, list = F)

wifi_train <- partTrainingData[in_training,]
wifi_test <- partTrainingData[-in_training,]
```

##### Pre-test with Principal Components Analysis
###### Classification
```{r}
# create train and test sets for pca
wifi_train_pca <- partTrainingData[in_training,]
wifi_test_pca <- partTrainingData[-in_training,]

# Remove columns with zero variance
rm_variables <- as.numeric(which(apply(wifi_train_pca, 2, var) == 0))
wifi_train_pca <- wifi_train_pca[-rm_variables]
wifi_test_pca <- wifi_test_pca[-rm_variables]
```

```{r}
pca_train <- wifi_train_pca %>% select(-ID)
pca_test <- wifi_test_pca %>% select(-ID)

pca <- prcomp(pca_train, scale = TRUE)

# variance
pr_var <- (pca$sdev)^2 

# % of variance
prop_varex <- pr_var/sum(pr_var)

# Plot
ggplot(data.frame(prop_varex), aes(x = 1:length(prop_varex), y = cumsum(prop_varex))) + 
  geom_point() + 
  labs(x = "Principal Component", 
       y = "Proportion of Variance Explained") +
  geom_hline(yintercept = 0.95, colour = "red", linetype = "dashed") +
  geom_vline(xintercept = 170, colour = "red", linetype = "dashed")
```

```{r, results='hide'}
rm(pr_var, rm_variables, prop_varex)
```

```{r}
# Creating a new dataset
new_train <- data.frame(ID = wifi_train_pca$ID, pca$x)
new_test <- data.frame(ID = wifi_test_pca$ID, predict(pca, newdata = pca_test))

new_train <- new_train[, 1:170]
new_test <-  new_test[, 1:170]
```

```{r}
# apply kNN algorithm
model_knn <- train(ID ~ ., data = new_train, method = "knn", 
                   tuneGrid = expand.grid(
                     k = 1:10
                   ),
                   trControl = fitControl)
   
pred_knn <- predict(model_knn, newdata = new_test)
postResample(new_test$ID, pred_knn)
```

```{r, results='hide'}
rm(pred_knn, model_knn, new_train, new_test, wifi_test_pca, wifi_train_pca, pca,
   pca_test, pca_train)
```

### Applied pre-build functions
#### Classification
```{r}
# methods <- c("knn", "C5.0", "ranger", "rf", "svmLinear", "svmRadial")
# 
# compare_methods <- c()
# 
# tic("Total")
# for(i in methods){
#   tic(paste("Method", i))
#   set.seed(1605)
#   model <- train(ID ~ ., data = wifi_train, method = i, 
#                  preProcess = c('zv', 'pca'),
#                  trControl = fitControl)
#   
#   pred <- predict(model, newdata = wifi_test)
#   pred_metric <- postResample(wifi_test$ID, pred)
#   compare_methods <- cbind(compare_methods, pred_metric)
#   toc()
# }
# toc()
# 
# colnames(compare_methods) <- methods

# saveRDS(compare_methods, "compare_methods.rds")
```

#### Regression
```{r}
# remove unneeded features prior to model fitting
longTrainingData <- trainingData %>% 
  select(starts_with("WAP"), LONGITUDE, FLOOR)

latTrainingData <- trainingData %>% 
  select(starts_with("WAP"), LATITUDE, FLOOR)
```

```{r}
# Randomly select - sample fixed fraction per group
set.seed(1605)
longTrainingData <- longTrainingData %>% sample_frac(size = 0.1)

set.seed(1605)
latTrainingData <- latTrainingData %>% sample_frac(size = 0.1)
```

##### Longitude
```{r}
#
# set.seed(1605)
# in_training <- createDataPartition(longTrainingData$LONGITUDE, p = 0.7, list = F)
# 
# wifi_train <- longTrainingData[in_training,]
# wifi_test <- longTrainingData[-in_training,]
```

```{r}
# methods <- c("knn", "rpart", "ranger", "rf", "svmLinear", "svmRadial", "gbm")
# 
# compare_long <- c()
# 
# tic("Total")
# for(i in methods){
#   tic(paste("Method", i))
#   set.seed(1605)
#   model <- train(LONGITUDE ~ ., data = wifi_train, method = i, 
#                  preProcess = c('center', 'scale', 'zv', 'pca'),
#                  trControl = fitControl)
#   
#   pred <- predict(model, newdata = wifi_test)
#   pred_metric <- postResample(wifi_test$LONGITUDE, pred)
#   compare_long <- cbind(compare_long, pred_metric)
#   toc()
# }
# toc()
# 
# colnames(compare_long) <- methods

# saveRDS(compare_long, "compare_long.rds")
```

##### Longitude
```{r}
#
set.seed(1605)
in_training <- createDataPartition(latTrainingData$LATITUDE, p = 0.7, list = F)

wifi_train <- latTrainingData[in_training,]
wifi_test <- latTrainingData[-in_training,]
```

```{r}
# methods <- c("knn", "rpart", "ranger", "svmLinear", "svmRadial", "gbm")
# 
# compare_lat <- c()
# 
# tic("Total")
# for(i in methods){
#   tic(paste("Method", i))
#   set.seed(1605)
#   model <- train(LATITUDE ~ ., data = wifi_train, method = i, 
#                  preProcess = c('center', 'scale', 'zv', 'pca'),
#                  trControl = fitControl)
#   
#   pred <- predict(model, newdata = wifi_test)
#   pred_metric <- postResample(wifi_test$LATITUDE, pred)
#   compare_lat <- cbind(compare_lat, pred_metric)
#   toc()
# }
# toc()
# 
# colnames(compare_lat) <- methods

# saveRDS(compare_lat, "compare_lat.rds")
```

#### Comparing metrics
```{r, results='asis'}
kable(compare_methods <- readRDS("compare_methods.rds"), digits = 3, 
      caption = "Classification model for buildingID + Floor")
```

```{r, results='asis'}
kable(compare_long <- readRDS("compare_long.rds"), digits = 3, 
      caption = "Regression models for longitude")
```

```{r, results='asis'}
kable(compare_lat <- readRDS("compare_lat.rds"), digits = 3, 
      caption = "Regression models for latitude")
```

#### Comparing time to process
```{r, results='asis'}
kable(as.data.frame(cbind("knn" = c(35.22), "C5.0" = c(512.06), 
                    "ranger" = c(389.30), "rf" = c(364.67), 
                    "smvLinear" = c(21.30), "smvRadial" = c(132.00))),
      digits = 2, caption = "Classification (time in seconds)")
```

```{r, results='asis'}
kable(as.data.frame(cbind("knn" = c(38.24), "rpart" = c(16.48), 
                    "ranger" = c(398.94), "rf" = c(1885.44), 
                    "smvLinear" = c(216.67), "smvRadial" = c(55.22), 
                    "gmb" = c(83.22))),
      digits = 2, caption = "Regression longitude (time in seconds)")
```

```{r, results='asis'}
kable(as.data.frame(cbind("knn" = c(42.41), "rpart" = c(16.08), 
                    "ranger" = c(406.76), "smvLinear" = c(212.99),
                    "smvRadial" = c(66.28), "gmb" = c(84.06))),
      digits = 2, caption = "Regression for latitude (time in seconds)")
```


### Final Models
#### Classification
```{r}
# remove unneeded features prior to model fitting
partTrainingData <- trainingData %>% 
  select(starts_with("WAP"), ID)
```

```{r}
# Randomly select - sample fixed fraction per group
set.seed(1605)
partTrainingData <- partTrainingData %>% sample_frac(size = 0.3, weight = ID)
```

```{r}
#
set.seed(1605)
in_training <- createDataPartition(partTrainingData$ID, p = 0.7, list = F)

wifi_train <- partTrainingData[in_training,]
wifi_test <- partTrainingData[-in_training,]
```

```{r}
# apply kNN algorithm
tic("kNN")
set.seed(1605)
model_knn <- train(ID ~ ., data = wifi_train, method = "knn", 
                   preProcess = c('zv', 'pca'),
                   tuneGrid = expand.grid(k = 1:10),
                   trControl = fitControl)
toc()

pred_knn <- predict(model_knn, newdata = wifi_test)
postResample(wifi_test$ID, pred_knn)
```


#### Regression
```{r}
# remove unneeded features prior to model fitting
longTrainingData <- trainingData %>% 
  select(starts_with("WAP"), LONGITUDE, FLOOR)

latTrainingData <- trainingData %>% 
  select(starts_with("WAP"), LATITUDE, FLOOR)
```

```{r}
# Randomly select - sample fixed fraction per group
set.seed(1605)
longTrainingData <- longTrainingData %>% sample_frac(size = 0.3)

set.seed(1605)
latTrainingData <- latTrainingData %>% sample_frac(size = 0.3)
```

##### Longitude
```{r}
#
set.seed(1605)
in_training <- createDataPartition(longTrainingData$LONGITUDE, p = 0.7, list = F)

wifi_train <- longTrainingData[in_training,]
wifi_test <- longTrainingData[-in_training,]
```

```{r}
tic("kNN")
set.seed(1605)
model <- train(LONGITUDE ~ ., data = wifi_train, method = "knn",
               preProcess = c('center', 'scale', 'zv', 'pca'),
               trControl = fitControl)
toc()

pred <- predict(model, newdata = wifi_test)
postResample(wifi_test$LONGITUDE, pred)


# saveRDS(compare_long, "compare_long.rds")
```

##### Latitude
```{r}
#
set.seed(1605)
in_training <- createDataPartition(latTrainingData$LATITUDE, p = 0.7, list = F)

wifi_train <- latTrainingData[in_training,]
wifi_test <- latTrainingData[-in_training,]
```

```{r}
tic("kNN")
set.seed(1605)
model_lat <- train(LATITUDE ~ ., data = wifi_train, method = "knn",
                   preProcess = c('center', 'scale', 'zv', 'pca'),
                   trControl = fitControl)
toc()

pred <- predict(model_lat, newdata = wifi_test)
postResample(wifi_test$LATITUDE, pred)

# saveRDS(compare_lat, "compare_lat.rds")
```

#### Validation
###### Convert data types
```{r}
# convert features to numeric
validationData <- sapply(validationData, as.numeric)
  
validationData <- as_tibble(validationData)
```

###### Impute NA's
```{r}
# convert NA's to -110
validationData[validationData == 100] <- -110
```

###### Remove features and Rows with all NA's
```{r}
# remove rows with all data is NA's
validationData <- validationData[rowSums(validationData == -110) != ncol(validationData), ]
```

###### Create unique identifier
```{r}
# consolidate position identifiers to create location ID feature
validationData$ID <- validationData %>%
  group_indices(BUILDINGID, FLOOR)
```

```{r}
# convert features to categorical
validationData <- validationData %>% 
  mutate(BUILDINGID = as.factor(BUILDINGID),
         SPACEID = as.factor(SPACEID),
         RELATIVEPOSITION = as.factor(RELATIVEPOSITION),
         FLOOR = as.factor(FLOOR),
         ID = as.factor(ID))
```

##### Classification
```{r}
# remove unneeded features prior to model fitting
classValidationData <- validationData %>% 
  select(starts_with("WAP"), ID)
```

```{r}
pred_class <- predict(model_knn, newdata = classValidationData)
postResample(classValidationData$ID, pred_class)
```


##### Regression
```{r}
# remove unneeded features prior to model fitting
longValidationData <- validationData %>% 
  select(starts_with("WAP"), LONGITUDE, FLOOR)

pred <- predict(model, newdata = longValidationData)
postResample(longValidationData$LONGITUDE, pred)

latValidationData <- validationData %>% 
  select(starts_with("WAP"), LATITUDE, FLOOR)

pred <- predict(model_lat, newdata = latValidationData)
postResample(latValidationData$LATITUDE, pred)
```



## Ensemble Models 

